Fluent statistical computing interfaces for biological data analysis

### 2020-03-19 Rmd/05-embeddings.Rmd
rebuild
add examples in...

There has been a rapid uptake in the use of non-linear dimensionality reduction (NLDR) methods such as t-distributed stochastic neighbour embedding (t-SNE) in the natural sciences as part of cluster orientation and dimension reduction workflows. The appropriate use of these methods is made difficult by their complex parameterisations and the multitude of decisions required to balance the preservation of
local and global structure in the resulting visualisation. We present a visual analytics framework for the pragmatic usage of NLDR methods by combining them with a technique called the tour. A tour is a sequence of interpolated linear projections of multivariate data onto a lower dimensional space. The sequence is displayed as a dynamic visualisation, allowing a user to see the shadows the high-dimensional data casts in a lower dimensional view. By linking the tour to a view
obtained from an NLDR method, we can preserve global structure and through user interactions like spatial linked brushing observe where
the NLDR view may be misleading. We display show several case 
studies from single cell genomics, that our approach is useful for
cluster orientation tasks. 
The implementation of our framework is available as an R package called `liminal` available at https://github.com/sa-lee/liminal.

@@

TODO

@@

@@

### Linking views via spatial brushes

@@

### Case Study 1: Fake trees

@@

>>>>

There has been a rapid uptake in the use of non-linear dimensionality reduction (NLDR) methods such as t-distributed stochastic neighbour embedding (t-SNE) in the natural sciences as part of cluster orientation and dimension reduction workflows. The appropriate use of these methods is made difficult by their complex parameterisations and the multitude of decisions required to balance the preservation of
local and global structure in the resulting visualisation. We present a visual analytics framework for the pragmatic usage of NLDR methods by combining them with a technique called the tour. A tour is a sequence of interpolated linear projections of multivariate data onto a lower dimensional space. The sequence is displayed as a dynamic visualisation, allowing a user to see the shadows the high-dimensional data casts in a lower dimensional view. By linking the tour to a view
obtained from an NLDR method, we can preserve global structure and through user interactions like spatial linked brushing observe where
the NLDR view may be misleading. We display several case 
studies from single cell genomics, that shows our approach is useful for
cluster orientation tasks. 
The implementation of our framework is available as an R package called `liminal` available at https://github.com/sa-lee/liminal.

@@

**distortion**: points that are far away in one view, will show neighbours that are close together in the opposing view.

**diffusion**: points that are close in one view, will show neighbours that are far apart in the opposing view.

@@

- **concatenated views:** see local and global structures
- **user interactions:** verify cluster shapes and locations
- **quality checks:** interrogate an embedding with reference
to the underlying data

@@

- don't need to realise the entire sequence
- instead generate new bases according to a fixed frame rate
- allows user interactions to play/pause on the current view

### Linking views via brushes

@@

### Case Study 1: Exploring tree structured data

(ref:fake-trees) Example from PHATE

```{r fake-trees, echo = FALSE, out.width = "50%", fig.cap="(ref:fake-trees)"}
knitr::include_graphics(here::here("img", "tree.png"))
```

#### Using t-SNE

- there's a lot of parameters to tweak
- emphasises locality, distance between clusters of points can be misleading
- size of clusters can be misleading 
- may require a few different runs to capture topology of data
- if there are clusters it will find them

@@

## Acknowledgements
